\chapter{Command Reference}
\label{app:cmdref}

\index{command reference}

All built-in commands, listed alphabetically.  Commands are executed by
middle-clicking (\button{2}) on the command name.

\begin{longtable}{@{}lp{3.6in}@{}}
\toprule
Command & Description \\
\midrule
\endhead
\cmd{Abort} &
  Debug command.  First execution prints a warning; second calls
  \texttt{abort()} to generate a core dump. \\
\cmd{Cut} &
  Delete selected text and copy to snarf buffer.  Marks window
  dirty. \\
\cmd{Del} &
  Close window.  Warns if dirty; execute again to force. \\
\cmd{Delcol} &
  Delete column and all its windows.  Warns if any are dirty. \\
\cmd{Delete} &
  Close window without dirty check. \\
\cmd{Dump} \textit{[file]} &
  Save session state.  Default: \file{\$HOME/acme.dump}. \\
\cmd{Edit} \textit{cmd} &
  Execute sam editing command on window body.  See
  Chapter~\ref{ch:edit}. \\
\cmd{Exit} &
  Quit acme.  Warns if any window is dirty; execute again to
  force. \\
\cmd{Font} \textit{[name]} &
  Toggle proportional/fixed font, or set font by name. \\
\cmd{Get} \textit{[file]} &
  Reload file from disk.  With argument, load named file. \\
\cmd{ID} &
  Print window ID number. \\
\cmd{Incl} \textit{[dirs...]} &
  List or add supplementary include directories for
  \texttt{<file.h>} resolution. \\
\cmd{Indent} \textit{[on|off|ON|OFF]} &
  Set auto-indent.  Lowercase: current window.  Uppercase: all
  windows. \\
\cmd{Kill} \textit{[names...]} &
  Send kill signal to named running commands. \\
\cmd{Load} \textit{[file]} &
  Restore session state.  Default: \file{\$HOME/acme.dump}. \\
\cmd{Look} \textit{[text]} &
  Literal text search in window body.  Uses selection or argument. \\
\cmd{New} \textit{[files...]} &
  Create new window.  With arguments, open named files. \\
\cmd{Newcol} &
  Create a new column. \\
\cmd{Paste} &
  Replace selection with snarf buffer contents. \\
\cmd{Put} \textit{[file]} &
  Write window to disk.  With argument, write to named file. \\
\cmd{Putall} &
  Write all dirty windows whose names are real files. \\
\cmd{Redo} &
  Redo last undone change. \\
\cmd{Send} &
  In Win windows: send pending input to shell.  In regular windows:
  append snarf buffer to body. \\
\cmd{Snarf} &
  Copy selected text to snarf buffer without deleting. \\
\cmd{Sort} &
  Sort windows in column alphabetically by name. \\
\cmd{Tab} \textit{[n]} &
  Print or set tab width for current window. \\
\cmd{Undo} &
  Undo last change or change group. \\
\cmd{Win} &
  Open interactive shell session in new window.  See
  Section~\ref{sec:win}. \\
\cmd{Zerox} &
  Clone window (second view of same file). \\
\bottomrule
\end{longtable}
