\chapter{Edit Command Reference}
\label{app:editref}

\index{Edit command!reference}

This appendix summarizes the sam editing language available via acme's
\cmd{Edit} command.  See Chapter~\ref{ch:edit} for tutorial coverage.

\section{Addresses}

\begin{longtable}{@{}lp{3.5in}@{}}
\toprule
Address & Meaning \\
\midrule
\endhead
\texttt{.} & Current selection (dot). \\
\texttt{0} & Beginning of file. \\
\texttt{\$} & End of file. \\
\texttt{\textit{n}} & Line \textit{n} (1-based). \\
\texttt{\#\textit{n}} & Character \textit{n} (0-based). \\
\texttt{/\textit{re}/} & Next match of regex \textit{re} forward. \\
\texttt{?\textit{re}?} & Next match of regex \textit{re} backward. \\
\texttt{\textit{a},\textit{b}} & Range from \textit{a} to \textit{b}. \\
\texttt{\textit{a};\textit{b}} & Range; set dot to \textit{a} before evaluating
  \textit{b}. \\
\texttt{\textit{a}+\textit{b}} & \textit{b} relative to end of \textit{a}. \\
\texttt{\textit{a}-\textit{b}} & \textit{b} relative to start of \textit{a},
  backward. \\
\texttt{,} & Shorthand for \texttt{0,\$} (entire file). \\
\bottomrule
\end{longtable}

\section{Commands}

\begin{longtable}{@{}lp{3.5in}@{}}
\toprule
Command & Action \\
\midrule
\endhead
\texttt{a/\textit{text}/} & Append \textit{text} after addressed range. \\
\texttt{c/\textit{text}/} & Change: replace addressed range with \textit{text}. \\
\texttt{d} & Delete addressed range. \\
\texttt{i/\textit{text}/} & Insert \textit{text} before addressed range. \\
\texttt{m \textit{addr}} & Move addressed range to after \textit{addr}. \\
\texttt{p} & Print addressed range. \\
\texttt{s/\textit{re}/\textit{text}/} & Substitute first match of \textit{re}. \\
\texttt{s/\textit{re}/\textit{text}/g} & Substitute all matches. \\
\texttt{t \textit{addr}} & Copy addressed range to after \textit{addr}. \\
\texttt{=} & Print address (line number and character offset). \\
\bottomrule
\end{longtable}

\section{Structural Commands}

\begin{longtable}{@{}lp{3.5in}@{}}
\toprule
Command & Action \\
\midrule
\endhead
\texttt{x/\textit{re}/\textit{cmd}} & Extract: for each match of \textit{re}, set
  dot to match and run \textit{cmd}. \\
\texttt{y/\textit{re}/\textit{cmd}} & Complement: for each stretch between
  matches, set dot and run \textit{cmd}. \\
\texttt{g/\textit{re}/\textit{cmd}} & Guard: run \textit{cmd} if addressed text
  matches \textit{re}. \\
\texttt{v/\textit{re}/\textit{cmd}} & Inverse guard: run \textit{cmd} if addressed
  text does \emph{not} match \textit{re}. \\
\texttt{\{ \textit{cmds} \}} & Group multiple commands. \\
\bottomrule
\end{longtable}

\section{Multi-File Commands}

\begin{longtable}{@{}lp{3.5in}@{}}
\toprule
Command & Action \\
\midrule
\endhead
\texttt{X/\textit{re}/\textit{cmd}} & For each file whose name matches
  \textit{re}, run \textit{cmd}. \\
\texttt{Y/\textit{re}/\textit{cmd}} & For each file whose name does \emph{not}
  match \textit{re}, run \textit{cmd}. \\
\bottomrule
\end{longtable}

\section{Substitute Backreferences}

In the replacement text of \texttt{s} commands:

\begin{longtable}{@{}lp{3.5in}@{}}
\toprule
Token & Meaning \\
\midrule
\endhead
\texttt{\&} & Entire match. \\
\texttt{\textbackslash 1} \ldots\ \texttt{\textbackslash 9} &
  Captured subexpressions. \\
\texttt{\textbackslash n} & Literal newline. \\
\texttt{\textbackslash t} & Literal tab. \\
\texttt{\textbackslash\textbackslash} & Literal backslash. \\
\bottomrule
\end{longtable}
