\chapter{Mouse Reference}
\label{app:mouseref}

\index{mouse!reference}

\section{Single-Button Actions}

\begin{longtable}{@{}llp{2.5in}@{}}
\toprule
Button & Location & Action \\
\midrule
\endhead
\button{1} & Body/Tag & Place cursor or select text. \\
\button{1} & Scrollbar & Scroll up (line at click moves to top). \\
\button{1} & Layout button & Drag to move window. \\
\button{1} double & Body/Tag & Select word. \\
\button{1} double & On bracket & Select to matching bracket. \\
\button{1} double & On quote & Select to matching quote. \\
\button{1} triple & Body/Tag & Select line. \\
\midrule
\button{2} & Body/Tag & Execute text as command. \\
\button{2} & Scrollbar & Jump scroll (proportional). \\
\midrule
\button{3} & Body/Tag & Open file, follow address, or search. \\
\button{3} & Scrollbar & Scroll down (top line moves to click). \\
\midrule
Scroll up & Body & Scroll text up. \\
Scroll down & Body & Scroll text down. \\
\bottomrule
\end{longtable}

\section{Chords}

\begin{longtable}{@{}lp{3.5in}@{}}
\toprule
Chord & Action \\
\midrule
\endhead
\chord{1}{2} & Cut: delete selection to snarf buffer. \\
\chord{1}{3} & Paste: replace selection with snarf buffer. \\
\chord{1}{2}+\button{3} & Snarf: copy selection to snarf buffer (no delete). \\
\button{2} hold, \button{1} click & 2-1 chord: pass \button{1} text as argument to
  \button{2} command. \\
\bottomrule
\end{longtable}

\section{Modifier-Click Emulation}

For mice with fewer than three buttons:

\begin{longtable}{@{}ll@{}}
\toprule
Modifier Click & Equivalent \\
\midrule
\endhead
\key{Ctrl}+\button{1} & \button{2} (execute) \\
\key{Alt}+\button{1} & \button{3} (look/open) \\
\bottomrule
\end{longtable}

\section{Button 3 Decision Table}

What happens when you right-click (\button{3}) on different kinds of
text:

\begin{longtable}{@{}lp{3.0in}@{}}
\toprule
Text & Action \\
\midrule
\endhead
\file{filename} & Open file (or jump to existing window). \\
\file{filename:N} & Open file at line N. \\
\file{filename:N.M} & Open file at line N, column M. \\
\file{filename:N.M,N.M} & Open file at line/column range. \\
\file{filename:/re/} & Open file, search for regex. \\
\texttt{:N} & Jump to line N in current file. \\
\texttt{:/re/} & Search for regex in current file. \\
\texttt{http://...} & Open URL in browser. \\
\texttt{<file.h>} & Search include directories. \\
Other text & Literal search in current body. \\
\bottomrule
\end{longtable}
