\chapter{Keyboard Reference}
\label{app:kbdref}

\index{keyboard!reference}

\section{Cursor Movement}

\begin{longtable}{@{}lp{3.2in}@{}}
\toprule
Key & Action \\
\midrule
\endhead
\key{Left Arrow} & One character left. \\
\key{Right Arrow} & One character right. \\
\key{Up Arrow} & One line up (sticky column; scrolls at edge). \\
\key{Down Arrow} & One line down (sticky column; scrolls at edge). \\
\key{Ctrl+Left} & Beginning of line. \\
\key{Ctrl+Right} & End of line. \\
\key{Ctrl+A} & Beginning of line. \\
\key{Ctrl+E} & End of line. \\
\bottomrule
\end{longtable}

\section{Page and File Navigation}

\begin{longtable}{@{}lp{3.2in}@{}}
\toprule
Key & Action \\
\midrule
\endhead
\key{Page Up} & Scroll up one screen. \\
\key{Page Down} & Scroll down one screen. \\
\key{Ctrl+Up} & Scroll up one screen. \\
\key{Ctrl+Down} & Scroll down one screen. \\
\key{Home} & Cursor to first visible character. \\
\key{End} & Cursor to start of last visible line. \\
\key{Ctrl+Home} & Beginning of file. \\
\key{Ctrl+End} & End of file. \\
\key{Ctrl+Page Up} & Beginning of file. \\
\key{Ctrl+Page Down} & End of file. \\
\bottomrule
\end{longtable}

\section{Editing}

\begin{longtable}{@{}lp{3.2in}@{}}
\toprule
Key & Action \\
\midrule
\endhead
\key{Backspace} & Delete character left. \\
\key{Delete} & Delete character right. \\
\key{Ctrl+H} & Delete character left (same as Backspace). \\
\key{Ctrl+W} & Delete word left. \\
\key{Ctrl+U} & Delete to beginning of line. \\
\key{Enter} & Insert newline (auto-indent if enabled). \\
\key{Tab} & Insert literal tab. \\
\key{Escape} & Select recent typed text (cut on second press). \\
\bottomrule
\end{longtable}

\section{Composition}

\begin{longtable}{@{}lp{3.2in}@{}}
\toprule
Key & Action \\
\midrule
\endhead
\key{Alt} (press and release) & Toggle composition mode. \\
Then type sequence & Insert composed Unicode character. \\
\key{Alt} again & Cancel composition. \\
\bottomrule
\end{longtable}

See Appendix~\ref{app:unicode} for composition sequences.
