\chapter*{Bibliography}
\addcontentsline{toc}{chapter}{Bibliography}
\label{ch:biblio}

\begin{description}[style=nextline,leftmargin=0pt,labelindent=0pt]

\item[Pike, Rob. ``Acme: A User Interface for Programmers.'']
Proceedings of the Winter 1994 USENIX Conference, San Francisco, 1994.\\
\url{https://research.swtch.com/acme.pdf}\\
The foundational paper describing acme's design, philosophy, and
interface.  Essential reading for understanding why acme works the way
it does.

\item[Pike, Rob. ``The Text Editor sam.'']
\emph{Software---Practice and Experience}, Vol.~17, No.~11,
pp.~813--845, November 1987.\\
\url{https://doc.cat-v.org/plan_9/4th_edition/papers/sam/}\\
The full paper on sam's design and command language, which acme's
\cmd{Edit} command implements.

\item[Pike, Rob. ``Structural Regular Expressions.'']
Proceedings of the EUUG Spring 1987 Conference, Helsinki, 1987.\\
\url{https://doc.cat-v.org/bell_labs/structural_regexps/}\\
Describes the theoretical foundation for sam's \texttt{x} and \texttt{y}
commands---applying regular expressions to structure rather than lines.

\item[Cox, Russ. ``A Tour of Acme'' (screencast).]
\url{https://research.swtch.com/acme}\\
\url{https://www.youtube.com/watch?v=dP1xVpMPn8M}\\
A video demonstration of day-to-day acme use by one of its primary
maintainers.  The best introduction for new users.

\item[acme(1) manual page.]
\url{https://9fans.github.io/plan9port/man/man1/acme.html}\\
The official manual page from plan9port.

\item[sam(1) manual page.]
\url{https://9fans.github.io/plan9port/man/man1/sam.html}\\
Reference for the sam command language used by acme's \cmd{Edit}
command.

\item[Pike, Rob. ``Plumbing and Other Utilities.'']
\url{https://doc.cat-v.org/plan_9/4th_edition/papers/plumb/}\\
Design and implementation of the plumber, Plan~9's inter-application
message routing system.

\item[Pike, Rob. ``A Minimalist Global User Interface.'']
\url{https://doc.cat-v.org/plan_9/4th_edition/papers/rio/}\\
Describes rio, Plan~9's window system, which shares acme's philosophy of
text and files as the universal interface.

\item[Plan 9 from User Space (plan9port).]
\url{https://github.com/9fans/plan9port}\\
The port of Plan~9 user-space tools to Unix, from which this standalone
acme is derived.

\item[Plan 9 Fourth Edition Papers.]
\url{https://doc.cat-v.org/plan_9/4th_edition/papers/}\\
The complete collection of Plan~9 documentation.

\item[Acme community resources.]
\url{https://acme.cat-v.org/}\\
Community-maintained collection of acme tips, scripts, and links.

\end{description}
